---
title: "AOI Fiat"
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, 
  fig.height = 4
)

library(AOI)
```

The `AOI` package provides a bridge between the spatial data stored in `rnaturalearth` and `USAboundaries` to provide consisent programatic access to fiat (non-physical boundaries that exclusively depend on acts of human decision) boudaries. 

Using the `aoi_get` query users can query country, USA state, and USA county data from a single programatic interface. Examples of these querys can be seen below. In all examples the `AOI` function `aoi_map` is used to generate quick `leaflet` views of the requested AOI(s)

# AOI by **country**

An AOI can be defined by a **country** name, abbrivaiation, or ISO code:

## Single

Contry data can be queried individually,

```{r, message = FALSE}
aoi_get(country = 'CR') %>% 
  aoi_map(returnMap = TRUE)
```

## Multiple

As a set,

```{r}
aoi_get(country = c('Brazil', 'PERU', "chile")) %>% 
  aoi_map(returnMap = TRUE)
```

## Regional

Or by regions...

```{r}
aoi_get(country = c("south asia", "africa")) %>% 
  aoi_map(returnMap = TRUE)
```

#  AOI by **state**  

An AOI in the United States,can be defined by **state** name or abbreviation.

## Single

Again single states,

```{r, message = FALSE}
aoi_get(state = 'CA') %>% aoi_map(returnMap = T)
```

## Multiple **states**:

```{r, message = FALSE}
aoi_get(state = c("Oregon", "Washington", "CA")) %>% aoi_map(returnMap = TRUE)
```

If you want the lower 48 USA state you can query `state = 'conus'`. 

```{r, message = FALSE}
aoi_get(state = "conus") %>% aoi_map(returnMap = TRUE)
```

## Region 

```{r, message = FALSE}
aoi_get(state = "south") %>% aoi_map(returnMap = TRUE)
```

## Combined calls

Calls to `country` can be mixed with calls to `state`.

```{r, message = FALSE}
aoi_get(country = "Mexico", state = c("CA", "NM", "AZ", 'TX')) %>% 
  aoi_map(returnMap = TRUE)
```

# AOI by **county** 

If you want to get the boundaries of a state with the county features intact, simply call aoi_get and set `county = 'all'`

```{r, message = FALSE}
aoi_get(state = "AL", county = 'all') %>% aoi_map(returnMap = TRUE)
```


## Single

```{r, message = FALSE}
aoi_get(state = "CO", county = "El Paso") %>% 
  aoi_map(returnMap = TRUE)
```

## Multiple

```{r, message = F}
aoi_get(state ="CA", county = c("Santa Barbara", "San Luis Obispo", "Ventura")) %>% aoi_map(returnMap = T)
```

## Counties in different states

```{r, message = F}
aoi_get(state =c("CA", "CO"), county = c("Santa Barbara", "El Paso")) %>% aoi_map(returnMap = T)
```

```{r, message = F}
aoi_get(state ='northeast', county = 'all') %>% 
  aoi_map(returnMap = T)
```

# Unioning

The union parameter automatically submits a call to st_union joining all returned features as a single geometry.

```{r, messages = FALSE}
aoi_get(state = c('CA', 'OR', 'WA'), union = TRUE) %>% 
  aoi_map(returnMap = TRUE)
```


```{r, messages = FALSE}
aoi_get(state = 'conus', union = TRUE) %>% 
  aoi_map(returnMap = TRUE)
```


```{r, messages = FALSE}
aoi_get(state = 'south', union = TRUE) %>% 
  aoi_map(returnMap = TRUE)
```

