---
title: "Bounding Boxes"
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  #comment = "#>",
  fig.width = 6, 
  fig.height = 4
)

library(AOI)
library(leaflet)
```

## BB coords - BB geom

When working with AOIs many operations can be simplified to working with bounding boxes. As such AOI offers expediate functions to get bounding boxes as geometries or coordinates. 

Here we run through the steps of finding a spatial geometry for the state of Virgina using `aoi_get` and then determine the resulting bounding box in both spatial and coordinate representations.

```{r}
# Get geometry for the state of Virgina
VA = aoi_get(state = "VA")
# Get the spatial bounding box
(bb = bbox_get(VA))

# Identify the bounding box coordinates from either object
(bb1 = bbox_coords(bb))
(bb2 = bbox_coords(VA))
(identical(bb1, bb2))
```

Often when working with AOIs there is a desire to modify the boundaries  for specific task. The AOI package offers a buffering function which simplifies the sf::st_buffer() function for this task While less flexable then st_buffer, it handles unit and projection defintions within the function and adds (or subtracts) a consistent disntace from all edges in either kilomters or miles (default). 

```{r}
AOI = list()
AOI[['normal']] = aoi_get("UCSB") 
AOI[['big']] = AOI$normal %>% aoi_buffer(1) 
AOI[['small']] = AOI$normal %>% aoi_buffer(-1, km = TRUE) 
```

```{r, echo=FALSE}

leaflet() %>% addTiles() %>% 
  addPolygons(data = AOI$big, 
              fillColor = "transparent", color = 'green',
              label = " AOI + 1 mile") %>% 
  addPolygons(data = AOI$normal, 
              fillColor = "transparent", color = 'blue',
              label = "AOI") %>% 
  addPolygons(data = AOI$small,
              fillColor = "transparent", color = 'red',
              label = "AOI - 1 KM")
```

## Is inside

Lastly, there 

```{r}

ca = aoi_get(state = "CA")
sb = aoi_get(state = "CA", county = "Santa Barbara")
(aoi_inside(ca, sb))

el_paso = aoi_get(state = "CO", county = "El Paso")
(aoi_inside(ca, el_paso))

```

```{r}

tahoe = aoi_get("lake tahoe")
aoi_inside(ca, tahoe, total = FALSE)
aoi_inside(ca, tahoe, total = TRUE)
```

