---
title: "Bounding Boxes"
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  #comment = "#>",
  fig.width = 6, 
  fig.height = 4
)

library(AOI)
library(leaflet)
```

## BB coords - BB geom

When working with AOIs many operations can be simplified to working with bounding boxes. As such AOI offers expediant functions to get bounding boxes as geometries or coordinates. 

Here we run through the steps of finding a spatial geometry for the state of Virginia and then determining the resulting bounding box in both spatial and coordinate representations.

```{r}
# Get geometry for the state of Virginia
VA = aoi_get(state = "VA")
# Get the spatial bounding box
(bb = bbox_get(VA))

# Identify the bounding box coordinates from either object
(bb1 = bbox_coords(bb))
(bb2 = bbox_coords(VA))
(identical(bb1, bb2))
```

Often when working with AOIs there is a desire to modify the boundaries for specific task. The AOI package offers a buffering function which simplifies the `sf::st_buffer()` function for this task. While less flexible then `st_buffer`, it handles unit and projection definitions within the function and adds (or subtracts) a consistent distance from all edges in either kilometers or miles (default). 

```{r}
AOI = list()
AOI$normal = aoi_get("UCSB") 
AOI$big = aoi_buffer(AOI$normal, 1) 
AOI$small = AOI$normal %>% aoi_buffer(-1, km = TRUE) 
```

```{r, echo=FALSE}
leaflet() %>% addTiles() %>% 
  addPolygons(data = AOI$big, 
              fillColor = "transparent", color = 'green',
              label = " AOI + 1 mile") %>% 
  addPolygons(data = AOI$normal, 
              fillColor = "transparent", color = 'blue',
              label = "AOI") %>% 
  addPolygons(data = AOI$small,
              fillColor = "transparent", color = 'red',
              label = "AOI - 1 KM")
```

## Is inside

Lastly, there often a desire to test if one boundary is within another as a test. This can often be done using sf::st_within or sf::st_intersects but requires carefull attention to paramter order and then a conversion of a sparse matrix to a boolean condition. Instead, `aoi_inside` checks if either of the two inputs are within

```{r}

ca = aoi_get(state = "CA")
sb = aoi_get(state = "CA", county = "Santa Barbara")

sf::st_within(sb, ca)
(aoi_inside(ca, sb))

el_paso = aoi_get(state = "CO", county = "El Paso")
(aoi_inside(ca, el_paso))

?aoi_inside

```


Comlpete or partical containment?

```{r}
tahoe = aoi_get("lake tahoe")
aoi_inside(ca, tahoe, total = FALSE)
aoi_inside(ca, tahoe, total = TRUE)
```

