---
title: "Example Use Cases"
description: |
  "Getting Fiat Data"
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output: distill::distill_article
vignette: >
  %\VignetteIndexEntry{fiat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dev = "jpeg"
)

gdalio_terra <- function(dsn, ..., band_output_type = "numeric") {
  v <- gdalio_data(dsn, ..., band_output_type  = band_output_type)
  g <- gdalio_get_default_grid()
  r <- terra::rast(terra::ext(g$extent), nrows = g$dimension[2], ncols = g$dimension[1], crs = g$projection)
  if (length(v) > 1) terra::nlyr(r) <- length(v)
  terra::setValues(r, do.call(cbind, v))
}
```

```{r setup}
library(AOI)
library(terra)
```


# `climateR` Integration

climateR provides access to spatial datasets available as web accessible resources. climateR utilities require an input AOI parameter that can be generated with AOI fucntions:

```{r}
library(climateR)

# fiat boundary request
AOI = aoi_get(state = "FL")
tmax = getGridMET(AOI = aoi_get(state = "FL"),
                     varname = "tmmx",
                     startDate = "2023-10-29")
```

```{r}
plot(tmax$daily_maximum_temperature)
```

```{r}
nlcd = getNLCD(AOI = aoi_ext("Fort Collins", wh = 10, units = "km", bbox = TRUE))

terra::plot(nlcd$`2019 Land Cover L48`)
```

# `zonal` Integration

zonal provides tools to summarized gridded data to polygon areas. Paired with climateR and AOI, it becomes easier to extract area based summaries of geopspatial landscape characteristics.

```{r}
library(zonal)

# fiat boundary request
AOI = aoi_get(state = "conus", county = "all")

tmax = getGridMET(AOI = AOI,
                  varname = "pr",
                  startDate = "2023-10-29")

s = execute_zonal(tmax, geom = AOI, ID = "fip_code", fun = "mean")

plot(s['mean.pr_2023.10.29'], border = NA)
```

# `gdalio` Integration 

`gdalio` is designed to read data direct with GDAL warp with the design philosophy that a target grid be specified once.  `AOI::aoi_ext` and `AOI::discritize` allow target grids to be easily defined.

```{r}
library(gdalio) ## https://github.com/hypertidy/gdalio

# Define target grid and discritization
(target = aoi_ext("Fort Collins", wh = 10, units = "km") |> 
  discritize(dim = c(1024))) 

# Define data source (taken from gdalio readme)
virtualearth_imagery <- tempfile(fileext = ".xml")

writeLines('<GDAL_WMS>
  <Service name="VirtualEarth">
    <ServerUrl>http://a${server_num}.ortho.tiles.virtualearth.net/tiles/a${quadkey}.jpeg?g=90</ServerUrl>
  </Service>
  <MaxConnections>4</MaxConnections>
  <Cache/>
</GDAL_WMS>', virtualearth_imagery)

# Set grid and read data:
gdalio_set_default_grid(target)
(img <- gdalio_terra(virtualearth_imagery, bands = 1:3))

plotRGB(img)

```

# `nhdplusTools` Integration 

`nhdplusTools` is a large-scale package designed for working with the NHDPlus data model. Additionally, it provides a suite of wrappers around the USGS web services to read data. 

```{r}
library(nhdplusTools)

x = get_nhdplus(AOI = aoi_ext(xy = c(-110,37), wh = 10000, bbox = TRUE), realization = "all")

{
  plot(x$catchment$geometry, main = "NHDPlus Access")
  plot(x$flowline$geometry, col = "blue", add = TRUE)
  plot(x$outlet, pch = 16, col = "red", add = TRUE)
}
```

# `osmdata` Integration 

OpenStreetMap is perhaps the ultimate example of open data and the `osmdata` package provides easy access to this resourece. The package forms Overpass Queries based on input bounding boxes like those produced by AOI.

```{r}
library(osmdata)

bb = aoi_ext("500 Linden Street, Fort Collins", wh = .5, units = "km")

roads <- opq(bbox = bb) %>%
    add_osm_feature(key = 'highway') %>%
    osmdata_sf () 

plot(roads$osm_lines['osm_id'])
```

# `FedData` Integration  

FedData is an ROpenSci package that facilitated data access to a range of products. Most requests are made using a `template` that defines the area to extract data for. 

```{r}
library(FedData) # https://docs.ropensci.org/FedData/articles/FedData.html

bb = aoi_ext("500 Linden Street, Fort Collins", wh = .5, units = "km", bbox =TRUE)

ssurgo <- get_ssurgo(
  template = bb,
  label = "foco"
)

plot(ssurgo$spatial["MUKEY"])

```


# Remote Vector Data

The FlatGeoBuff team hosts dataset of every census block in the USA, including its population. The entire file is over 12GB - this remote file can be subset using the GDAL/sf with a passed AOI.

Again, the magic is being done by the ogr2og/GDAL, while AOI provides the convinence of easily defined extents and areas.

```{r}
library(sf)

bb = aoi_ext("500 Linden Street, Fort Collins", wh = 2, units = "km", bbox = TRUE) 

xx = read_sf('/vsicurl/https://flatgeobuf.septima.dk/population_areas.fgb', wkt_filter = st_as_text(bb))
```

```{r}
plot(xx['population'])
```

# Remote Raster Data (variable CRS)

A 100+ GB, tiled DEM is hosted by the USGS as part of the National Map. In this same bucket, a VRT file was built which can be read with `terra::rast()`. Simular to how sf::read_sf allows for a partical file read using the wkt_filter, 

```{r}
r = rast('/vsicurl/https://prd-tnm.s3.amazonaws.com/StagedProducts/Elevation/1/TIFF/USGS_Seamless_DEM_1.vrt')

bb = aoi_ext("500 Linden Street, Fort Collins", wh = .5, units = "km", bbox = TRUE, crs = crs(r))

d = crop(r, bb)
```

```{r}
plot(d)
```
