---
title: "Fiat Boundaries"
description: |
  "Getting Fiat Data"
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output: distill::distill_article
vignette: >
  %\VignetteIndexEntry{fiat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dev = "jpeg"
)
```

```{r setup}
library(AOI)
library(mapview)
library(dplyr)
```

AOI aims to provide a consistent, and flexible, syntax for retrieving fiat boundaries for spatial analysis, map making, and data retrivial. This is all done in the `aoi_get` function:

> **Definition**: _fiat_: fiatness implies mind-dependence, and thus a fiat boundary is a boundary that is determined by human fiat, lacking any natural indication.

This package makes no claims on sovereignty and returns data as found in Natural Earth, or, produced by the US Census Department (in the US).

# Countries

To retrive country boundaries (1 or more), ISO codes, abbreviations and names can be used:

```{r}
am = aoi_get(country = "New Zealand") |> 
  plot(max.plot = 1)

aoi_get(country = c("484", "USA", "CA")) |> 
  plot(max.plot = 1)
```

# State

```{r}
aoi_get(state = "CO") |> 
  plot(max.plot = 1)

aoi_get(state = c("CO", "Utah")) |> 
  plot(max.plot = 1)
```

# County


```{r}
aoi_get(state = "CO", county = "Larimer") |> 
  plot(max.plot = 1)

aoi_get(state = "TX", county = "all") |> 
  plot(max.plot = 1)
```

# FIP

```{r}
aoi_get(fip = "01")  |> 
  plot(max.plot = 1)

aoi_get(fip = "01125")  |> 
  plot(max.plot = 1)
```

# Zipcode

```{r}
aoi_get(zipcode = "80906") |> 
  plot(max.plot = 1)
```

# Aggregate
 
```{r}
aoi_get(country = "Asia") |> 
  plot(max.plot = 1)

aoi_get(country = "Latin America & Caribbean") |> 
  plot(max.plot = 1)

aoi_get(state = "conus") |> 
  plot(max.plot = 1)

aoi_get(state = "south") |> 
  plot(max.plot = 1)
```

# Union

```{r}
aoi_get(country = "Latin America & Caribbean", union = TRUE) |> 
  plot()
```
