---
title: "Geocoding"
description: |
  "Geocoding and spatial representation"
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output: distill::distill_article
vignette: >
  %\VignetteIndexEntry{Geocoding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dev = "jpeg"
)

library(dplyr)
```

```{r setup}
library(AOI)
```

Geocoding is the act of converting a named entity to a spatial representation. `AOI` uses the [`tidygeocoder`](https://jessecambon.github.io/tidygeocoder/) package for geocoding and overlays some basic syntax to return optional representations.

The `tidygeocoder` package provides a range of services (see ?tidygeocoder::geo mehtods). You are free to chose your own with the `method` argument. The default method for the package is `arcgis`.

```{r, default-method}
(AOI:::default_method)
```

## Basic Usage

The default `AOI::geocode` behavior returns a` data.frame` with a "request" named after the input `geo`, and an `x`, `y` column along with the `{method}_address` and `score`.

```{r, basic-usage}
geocode("Fort Collins") |>
  glimpse()
```

## Return a POINT representation

The basic return can be converted to a POINT representation by setting `pt = TRUE`. 

```{r}
geocode("Fort Collins", pt = TRUE) |> 
  glimpse()
```

The data will be returned in the crs specified in the function signature. By default, all AOI utilities act on:

This can be changed to meet your application needs: 

```{r}
geocode("Fort Collins", pt = TRUE, crs = 5070) |> 
  glimpse()
```
`
AOI::geocode` is setup to take more then one request and all prior arguments apply:

```{r}
geocode(c("Boulder", "Fort Collins"), pt = TRUE, crs = 5070) |> 
  glimpse()
```


## Return bounding box representation

Setting `bbox = TRUE` will return the bounding box representation of the `geo` input using the following rules:

1. If the geocoding method does **not** return an extent for an input, the POINT is returned
2. If a single `geo` is passed _AND_ and extent is returned, the `POLYGON` is generated
3. If multiple `geo` objects are passed, the POLYGON representing the extent of all POINT(s) is returned

### Default behavior

```{r}
geocode("Fort Collins", bbox = TRUE)  |> 
  glimpse()
```

### Multiple inputs return a single bounding box:

```{r}
# EPSG:5070
geocode(c("Boulder", "Fort Collins"), bbox = TRUE, crs = 5070)  |> 
  aoi_map(returnMap = TRUE)
```

If you actually _do_ want the extents of multiple inputs, then pass the inputs through `lapply`:

```{r}
lapply(c('Fort Collins', 'Boulder'), FUN = geocode, bbox = TRUE) |>
  bind_rows() |>
  aoi_map(returnMap = TRUE)
```


## Return bounding box and POINT representations

Setting `all = TRUE` will return both the `point` and `bbox` representation. The `bbox` returned will still follow the same rules:

```{r}
geocode('Fort Collins', all = TRUE) |>
  aoi_map(returnMap = TRUE)

geocode(geo = c('Fort Collins', 'Boulder'), all = TRUE) |>
  aoi_map(returnMap = TRUE)
```


## xy

Setting `xy = TRUE` will return just the XY coordinates of the geocoded point. This is useful for applications shown later:

```{r}
geocode("Fort Collins", xy = TRUE)
geocode(geo = "Fort Collins", xy = TRUE, crs = 5070)
```

## Event based 

Moving away for just tidygeocdoing, AOI::geocode provides tha

```{r}
geocode(event = "dday") |>
 aoi_map(returnMap = TRUE)
```

```{r}
geocode(event = "Hurricane Harvey", all = TRUE) |>
  aoi_map(returnMap = TRUE)
```


# Reverse Geocoding

Reverse geocoding is the act of converting a XY location to a named entity. Here we work through an example of a place 

```{r}
# getting an xy input
xy = geocode("UCSB", xy = TRUE)
```

```{r}
geocode_rev(xy)
```

```{r}
geocode_rev(xy, pt = TRUE)
```
