---
title: "Geocoding"
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, 
  fig.height = 4
)

devtools::load_all()
library(kableExtra)
library(leaflet)
library(sf)

```

The AOI package provides an interface to the Open Street Map [Nominatim API](https://nominatim.org/). This service does not require an API key and can be used for most research grade purposes. This page walks through a few examples of `AOI` geocoding:

# Basic Geocoding {.tabset .tabset-fade}

## By Place Name 

Basic geocoding functionality converts a place name into a lat/long coordinate pair.

```{r, eval = F}
geocode("UCSB") 
```
```{r, echo = FALSE}
geocode("UCSB") %>% 
  kable() %>% 
  kable_styling()
```

## By zip code (USA)

The AOI package provides basic geocoding by USA zip code by connecting with Tom Boutells archive of [zipcode centroids](https://www.boutell.com/zipcodes/).

```{r, eval = F}
geocode(zipcode = "80906")
```

```{r, echo = F}
geocode(zipcode = "80906") %>%
  kable() %>% 
  kable_styling()
```

## By Event (beta)

Sometimes we want to geocode something that happen but that isn't a set place that would appear in the OSM gazetteer. In these case AOI offers a beta tool to geocode by scrapping Wikipedia.

```{r, eval = F}
x = lapply(c("Deepwater Horizion", 
             "JFK Assasination", 
             "I have a dream speech"),
           geocode_wiki)

do.call(rbind, x)
```
```{r, echo = F}
x = lapply(c("Deepwater Horizion", 
             "JFK Assasination", 
             "I have a dream speech"),
           geocode_wiki)

do.call(rbind,x) %>% 
  kable() %>% 
  kable_styling()
```

## Reverse Geocoding

The oppisite of forward geocodeing is reverse geocodeing (going from lat/lon to place name). AOI offeres a reverse geocoding service that integrates results from ESRIs and OSMs APIs.

```{r, eval = FALSE}
revgeocode(c(37, -119))
```

```{r, echo = F}
revgeocode(c(37, -119)) %>% 
  t() %>% 
  kable() %>% 
  kable_styling() 
```


# Geocoding Extensions {.tabset .tabset-fade}

As a tool to support spatial research, AOI offers the abilty to `geocode()` multiple entries at a time and return spatial representations of the results:

## Geocode Multiple Entries

```{r, eval = F}
geocode(c("UCSB", "Goleta", "State Street, Santa Barbara")) 
```

```{r, echo = F}
geocode(c("UCSB", "Goleta", "State Street, Santa Barbara")) %>% 
  kable() %>% 
  kable_styling()
```

## Return Points

```{r}
do.call("rbind",
        lapply(c("UCSB", "Goleta", "State Street, Santa Barbara"), 
               function(x) geocode(x, pt= T)) ) %>% check(returnMap = T)

```

## Return Bounding Box

Setting `bb = TRUE` will return the OSM defined bounding box of the queried place name:

```{r}

geocode("Garden of the Gods", bb = T) %>% 
  check(returnMap = T)

```
 
## Return all

If you want the coordinate, point, and bounding box realizations of a query, set `all=TRUE`
 
```{r}

call = geocode(location = c("Garden of the Gods", "Quail Lake, Colorado"), all = T) 

str(call, max.level = 1)

check(call, returnMap = T)

```
 
## Turn on full return

By default geocoding returns only provides the lat/lon coordinates from the queried location. To recieve all information about a query, set `full = TRUE`

```{r, eval  =F }
geocode("Goleta", full= TRUE)
```

```{r, echo = F}
geocode("Goleta", full= TRUE) %>% t() %>% kable %>% kable_styling()
```
