---
title: "Extents & discritizations"
description: |
  "Defining Regions"
author:
  - name: "Mike Johnson"
    url: https://github.com/mikejohnson51
    affiliation: Lynker, NOAA-Affiliate
    affiliation_url: https://lynker.com
output: distill::distill_article
vignette: >
  %\VignetteIndexEntry{extents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dev = "jpeg"
)
```

```{r setup}
library(AOI)
library(dplyr)
```

## Basic Usage

The core functions described here are `aoi_ext` which builds a specific extent around a given `POINT` / `coordinate set` and `discritize` which discritizes an extent/object. 

To start, lets build a 1000 meter extent around the location called "Fort Collins":

```{r}
aoi_ext(geo = "Fort Collins", wh = 1000)
```
From this, we can see a extent was created!

## Add bbox

As with `AOI::geocode`, the returned object can be represented as a `POLYGON` by adding `bbox = TRUE`. This is helpful for mapping:

```{r}
aoi_ext(geo = "Fort Collins", wh = 1000, bbox = TRUE) %>% 
  mapview::mapview()
```

## Modify both diminisons

Providing a single value to the `wh` argument generates a square extent. To find a rectangular extent, a two value vector can be passed:

```{r}
aoi_ext(geo = "Fort Collins", wh = c(2000, 1000), bbox = TRUE) %>% 
  mapview::mapview()
```

## Use XY

We have relied on providing a place that requires `AOI::geocode` to be run. Instead, we can provide a direct coordinate set as an `xy vector`:

```{r}
aoi_ext(xy = c(x = -119, y = 36), wh = 1000)
```

## Units

By default all measures are made in the default AOI unit:

```{r}
(AOI:::default_units)
```

We can override these by defining the `units` argument in the aoi_ext signiture. Here we see that a 1km and 1000m wh argument produce identical results:

```{r}
identical( aoi_ext(geo = "Fort Collins", wh = 1, units = "km"), 
           aoi_ext(geo = "Fort Collins", wh = 1000) )
```

## Area discritizations

In many applications, getting the extent is only part of the challenge. The other is defining a discretizion of a region. This can be done with the `AOI::discritize` function:

```{r}
aoi_ext("Fort Collins", wh = c(10000, 20000)) %>% 
  discritize()
```

## Diminsions

By default, the function uses the AOI default discritization:

```{r}
(AOI:::default_dim)
```

This can be changed using the dim argument: 

```{r}
aoi_ext("Fort Collins", wh = c(10000, 20000)) %>% 
  discritize(dim = 1024)
```

## CRS

The default expected input CRS is EPSG:4326. This is also the default output. Both can be changed depending on the desired input and output.

 - If a CRS can be extracted from the input, then it is used.
 - If a CRS cannot be extracted from the input, then EPSG:4326 is assumed
 - The assumed CRS can be overridden with the  `in_crs` argument

```{r}
# CRS inferred from input and changed to 5070
aoi_ext("Fort Collins", wh = c(20000, 10000)) %>% 
  discritize(out_crs = 5070)
```

```{r}
# CRS inferred
t1 = aoi_ext("Fort Collins", wh = c(20000, 10000), crs = 5070, bbox = TRUE) %>% 
  discritize(out_crs = 5070)
t1$extent

# Bad, assumed EPSG
t2 = c(-781506.2, 1976590.7, -740107.0, 2000454.0) %>% 
  discritize(out_crs = 5070)
t2$extent

# Explicit CRS named
t3 = c(-781506.2, 1976590.7, -740107.0, 2000454.0)  %>% 
  discritize(in_crs = 5070, out_crs = 5070)
t3$extent
```


## SpatRaster

If you want a gridded object for future processing, you can set `spatrast = TRUE`.

```{r}
aoi_ext("Fort Collins", wh = c(20000, 10000)) %>% 
  discritize(spatrast = TRUE)
```

# Passing Other objects.

Defining and extent for a `POINT` and `wh` argument is only one option that can be discritized. Others include all AOI outputs, sf objects and both SpatVect and SpatRast objects.

## AOI BBOX returns

```{r}
geocode(c("Fort Collins", "Boulder"), bbox = TRUE) %>% 
  discritize(spatrast = TRUE, dim = c(10,10))
```

## AOI fiat boundaries

(really this applies to any sf object!)

```{r}
aoi_get(state = "OR") %>% 
  discritize(spatrast = TRUE, dim = c(5,5))
```

## Other `Spat*` object

```{r}
(r <- terra::rast(system.file("ex/elev.tif", package="terra")) )

# SpatRast
discritize(r, spatrast = TRUE)

# SpatVect
terra::vect(system.file("ex/lux.shp", package="terra")) %>% 
  discritize(spatrast = TRUE)
```
